### 2015-05-13 background.tex
addressed some comments

We believe that our focus in locks is justified, because the majority of Linux drivers still use locks. To check this, we used a simple grep command and found that 68\% of the drivers from the 4.0 Linux kernel distribution use locks, versus 13\% that use atomic operations. We also found that 31\% of the drivers use neither locking nor atomic primitives, which suggests that they are either simple (non-concurrent) drivers or use some other lock-free synchronization primitive.

>>>>

\PDComment{Do we really need the following? I am not even sure how entirely accurate it is as I just used some simple greping ...}
We believe that our focus in locks is justified, because the majority of Linux drivers still use locks. To check this, we used a simple grep command and found that 68\% of the drivers from the 4.0 Linux kernel distribution use locks, versus 13\% that use atomic operations. We also found that 31\% of the drivers use neither locking nor atomic primitives, which suggests that they are either simple (non-concurrent) drivers or use some other lock-free synchronization primitive.

### 2015-05-13 whoop.tex
addressed some comments

\medskip\noindent\textbf{Extracting Entry Point Information }
%
Chauffeur is a Clang-based frontend that performs two important tasks. The first is to traverse (using a Clang visitor pass) the abstract syntax tree (AST) of the original driver and extract all entry point identifier names, together with the identifier names of their corresponding kernel API functions. Linux drivers define entry points in a standard way (see Figure~\ref{fig:entrypoints} for an example of how the generic\_nvram driver defines the entry points for the \texttt{file\_operations} API); Chauffeur identifies these definitions in the AST and outputs the relevant information in an XML file, which is subsequently parsed by \whoop.

@@

\ADComment{This is not clear to me.  Is the purpose of the main function simply to make sure that the functions are not compiled away?  I.e., do we then just ignore the main function?  If that is the case then why does it matter what we pass to the entry points, i.e. why should this be some arbitrary value?}  The second task is to instrument the original driver with a function that calls all entry points, one after the other, passing them parameters of arbitrary value (to over-approximate kernel behavior for soundness). This function effectively acts as the \emph{main} function of the driver and is required so that Clang does not compile away any entry points that are not otherwise invoked (because we closed the environment using stub functions).  \ADComment{It is not clear to me why an entry point would be invoked anywhere in the code other than in the main function we add.  Is it the case that sometimes entry points call each other?  Or does the kernel sometimes make direct calls to named entry points (rather than calls via function pointers)?} This simple rewriting does not alter the semantics of the original driver, as \whoop only uses this function to create checker functions and instrument them with calls to entry points (see Figure~\ref{fig:sequentialization} for the pseudocode of such checker function). Chauffeur outputs the rewritten driver in a C file.

\medskip\noindent\textbf{Translation for Verification }
%
The rewritten driver is compiled by Clang to LLVM-IR~\cite{lattner2004llvm}, a low-level assembly-like language in single static assignment (SSA) form. Function calls (e.g.\ for locking and unlocking) are preserved in this translation, which is useful for our lockset instrumentation.

@@

\medskip\noindent\textbf{Verification and Error Reporting }
%
For each entry point pair, the instrumented sequential program, equipped with procedure and loop summaries, sent to the Boogie verification engine.  For each procedure in the program, Boogie generates a verification condition~\cite{barnett2005weakest} and discharges these to the Z3~\cite{de2008z3} theorem prover.  In particular, the verification for the root-level procedure, encoding the sequential program sketched in Figure~\ref{fig:sequentialization}, encodes the race-freedom check for the entry point pair.
Successful verification implies that the entry point pair is free of data races, while an error denotes a \emph{potential} data race and is reported to the user. To improve usability, \whoop has a built-in error reporter that matches counterexamples to source code. The following is an example of a reported error: \ADComment{Can you replace this with an error Whoop would give for the example device driver bug used earlier in the paper, and refer back to that example?  That would look more convincing.}

@@

First, we enriched \whoop with information regarding \emph{kernel-imposed serialization}, to increase precision. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized. This is an ongoing manual effort: the more drivers we study, the more such properties we discover, which we can then exploit to make \whoop more precise.  \ADComment{Can you make it clearer that this information is currently provided manually to Whoop?  Can you explain how this is achieved?  And can you comment briefly on how we figured this out?}

>>>>

\medskip\noindent\textbf{Extracting Entry Point Information }
%
Chauffeur is a Clang-based frontend that traverses the abstract syntax tree (AST) of the original driver (using a Clang visitor pass) and extracts all entry point identifier names, together with the identifier names of their corresponding kernel API functions. Linux drivers define entry points in a standard way (see Figure~\ref{fig:entrypoints} for an example of how the generic\_nvram driver defines the entry points for the \texttt{file\_operations} API); Chauffeur identifies these definitions in the AST and outputs the relevant information in an XML file, which is subsequently parsed by \whoop and used during the instrumentation.

@@

\medskip\noindent\textbf{Translation for Verification }
%
Next, the driver is compiled by Clang to LLVM-IR~\cite{lattner2004llvm}, a low-level assembly-like language in single static assignment (SSA) form. Function calls (e.g.\ for locking and unlocking) are preserved in this translation, which is useful for our lockset instrumentation.

@@

\medskip\noindent\textbf{Verification and Error Reporting }
%
For each entry point pair, the instrumented sequential program, equipped with procedure and loop summaries, sent to the Boogie verification engine.  For each procedure in the program, Boogie generates a verification condition~\cite{barnett2005weakest} and discharges these to the Z3~\cite{de2008z3} theorem prover.  In particular, the verification for the root-level procedure, encoding the sequential program sketched in Figure~\ref{fig:sequentialization}, encodes the race-freedom check for the entry point pair.
Successful verification implies that the entry point pair is free of data races, while an error denotes a \emph{potential} data race and is reported to the user. To improve usability, \whoop has a built-in error reporter that matches counterexamples to source code. The following is an example of a reported error: \ADComment{Can you replace this with an error \whoop would give for the example device driver bug used earlier in the paper, and refer back to that example?  That would look more convincing.}

@@

First, we enriched \whoop with information regarding \emph{kernel-imposed serialization}, to increase precision. The Linux kernel can serialize calls to entry points, thus forcing them to run sequentially with each other. As an example, a large number of networking entry points are mutually serialized with RTNL, a network-specific kernel lock. \whoop exploits this knowledge and does not create pairs for entry points that are serialized. This is an ongoing manual effort: the more drivers we study, the more such properties we discover, which we can then exploit to make \whoop more precise.  \ADComment{Can you make it clearer that this information is currently provided manually to \whoop?  Can you explain how this is achieved?  And can you comment briefly on how we figured this out?}
